Observations from Parallelising Three Maximum Common (Connected) Subgraph Algorithms

### 2017-08-09 paper/paper.tex
Move
Thing

?? Explain induced, give examples.

Finding a maximum common subgraph is the key step in measuring the similarity or difference between
two graphs \citep{DBLP:journals/prl/Bunke97,DBLP:journals/prl/FernandezV01,o:Kriege15}: to determine
the difference between two graphs, we find what they have in common, and then take everything left
over. Because of this, maximum common subgraph problems frequently arise in biology and chemistry
\citep{DBLP:journals/jcamd/RaymondW02a,o:EhrlichR11,DBLP:journals/dam/GayFMSS14} where graphs
represent molecules, but also in applications including computer vision
\citep{DBLP:journals/jair/CookH94,DBLP:conf/gbrpr/CombierDS13}, computer-aided manufacturing
\citep{o:LuoWSN17}, crisis management \citep{o:DelavalladeFLL16}, deanonymising datasets
\citep{o:SharadD13}, the analysis of source code \citep{DBLP:journals/tkde/DjokoCH97}, binary
programs \citep{DBLP:conf/icics/GaoRS08}, and circuit designs \citep{DBLP:journals/jair/CookH94}, in
malware detection \citep{DBLP:journals/compsec/ParkRS13}, social network analysis
\citep{DBLP:journals/tkde/FangYZZ15}, graph database query explanations
\citep{DBLP:journals/jcss/VasilyevaTBL16}, and in character recognition problems
\citep{DBLP:journals/pr/LuRS91}.

?? Connected, which occurs in many applications
\citep{DBLP:journals/tcs/Koch01,DBLP:journals/jcamd/RaymondW02a,DBLP:conf/mco/VismaraV08,o:EhrlichR11,o:LuoWSN17}.
Adding the connectedness requirement makes certain special cases solvable in polynomial time,
including outerplanar graphs of bounded degree \citep{DBLP:journals/algorithms/AkutsuT13} and trees
\citep{DBLP:conf/mfcs/DroschinskyKM16}, but the general case remains NP-hard. As illustrated in
\cref{figure:mcsexample}, the maximum common connected subgraph cannot be deduced from the maximum
common subgraph.

This paper looks at three recent branch and bound algorithms, each of which is the state of the art
for certain classes of instance.

The ultimate aim of this chapter is to parallelise a state-of-the-art algorithm for maximum common
subgraph problems, and to see whether we agree with \citepos{DBLP:conf/ictai/MinotNS15} conclusion
that doing so is unusually challenging.

@@

The maximum common subgraph problem may be reformulated as a constraint optimisation problem, as
follows. We pick whichever input graph has fewer vertices, and call it the \emph{pattern}; the other
graph is called the \emph{target}. The model is then to construct an injective, partial mapping from
the pattern to the target, preserving both adjacency and non-adjacency. Thus, for each
vertex in the pattern, we create a variable, whose domain ranges over each vertex in the target
graph, plus an additional value $\bot$. We then have three sets of constraints. The first set says
that for each pair of adjacent vertices in the pattern (that is, for each edge in the pattern),
if neither of these vertices are mapped to $\bot$ then these vertices must be mapped to an adjacent
pair of target vertices. The second set is similar, but looks at non-adjacent pairs (or non-edges).
Finally, the third set ensures injectivity, by enforcing that the variables must be all different
except when using $\bot$.

>>>>

A \emph{subgraph isomorphism} is an injective mapping from a \emph{pattern} graph to a \emph{target}
graph which maps adjacent vertices to adjacent vertices---that is, it \emph{preserves adjacency}. It
is \emph{induced} if additionally it maps non-adjacent vertices to non-adjacent vertices, preserving
non-adjacency. When working with labelled graphs, a subgraph isomorphism must preserve labels, and
on directed graphs, it must preserve orientation. A \emph{common induced subgraph} of two graphs $G$
and $H$ is a pair of induced subgraph isomorphisms from a pattern graph $P$, one to $G$ and one to
$H$. A \emph{maximum common induced subgraph} is one with as many vertices as possible.

Finding a maximum common subgraph is the key step in measuring the similarity or difference between
two graphs \citep{DBLP:journals/prl/Bunke97,DBLP:journals/prl/FernandezV01,o:Kriege15}: to determine
the difference between two graphs, we find what they have in common, and then take everything left
over. Because of this, maximum common subgraph problems frequently arise in biology and chemistry
\citep{DBLP:journals/jcamd/RaymondW02a,o:EhrlichR11,DBLP:journals/dam/GayFMSS14} where graphs
represent molecules, but also in applications including computer vision
\citep{DBLP:journals/jair/CookH94,DBLP:conf/gbrpr/CombierDS13}, computer-aided manufacturing
\citep{o:LuoWSN17}, crisis management \citep{o:DelavalladeFLL16}, deanonymising datasets
\citep{o:SharadD13}, the analysis of source code \citep{DBLP:journals/tkde/DjokoCH97}, binary
programs \citep{DBLP:conf/icics/GaoRS08}, and circuit designs \citep{DBLP:journals/jair/CookH94}, in
malware detection \citep{DBLP:journals/compsec/ParkRS13}, social network analysis
\citep{DBLP:journals/tkde/FangYZZ15}, graph database query explanations
\citep{DBLP:journals/jcss/VasilyevaTBL16}, and in character recognition problems
\citep{DBLP:journals/pr/LuRS91}.  A common variant of the problem requires a largest
\emph{connected} subgraph
\citep{DBLP:journals/tcs/Koch01,DBLP:journals/jcamd/RaymondW02a,DBLP:conf/mco/VismaraV08,o:EhrlichR11,o:LuoWSN17}.
As illustrated in \cref{figure:mcsexample}, the maximum common connected subgraph cannot be deduced
from the maximum common subgraph.

Although both variants are NP-hard, some progress has been made towards solving the problem in
practice.  This paper looks at three recent branch and bound algorithms for maximum common
(connected) induced subgraph problems, each of which is the state of the art for certain classes of
instance. We discuss our experiences in adding parallel tree-search to these three algorithms. In
each case, our results show that the parallel version of the algorithm is clearly better than the
sequential version, although a closer look at the results shows many nuances.

@@

\subsection{Constraint Programming}

The maximum common subgraph problem may be reformulated as a constraint optimisation problem, as
follows.  Observe that an equivalent definition of a common subgraph of graphs $G$ and $H$ is an
injective partial mapping from $G$ to $H$ which preserves both adjacency and non-adjacency. Hence we
pick whichever input graph has fewer vertices, and call it the \emph{pattern}; the other graph is
called the \emph{target}. The model is then to construct an injective, partial mapping from the
pattern to the target, preserving both adjacency and non-adjacency. Thus, for each vertex in the
pattern, we create a variable, whose domain ranges over each vertex in the target graph, plus an
additional value $\bot$. We then have three sets of constraints. The first set says that for each
pair of adjacent vertices in the pattern (that is, for each edge in the pattern), if neither of
these vertices are mapped to $\bot$ then these vertices must be mapped to an adjacent pair of target
vertices. The second set is similar, but looks at non-adjacent pairs (or non-edges).  Finally, the
third set ensures injectivity, by enforcing that the variables must be all different except when
using $\bot$.
